# Reproducible research - Peer Assessment #1
Prepared by John Romankiewicz, Berkeley, CA
Date: 10/15/2015

First you'll need to clone the repository at this link:
http://github.com/rdpeng/RepData_PeerAssessment1

```{r}
temp <- unzip("activity.zip")
data <- read.csv(temp, header=T)
```

## Question 1: What is mean total number of steps taken per day?

Below, we will calculate the mean and median number of steps taken per day, as well as provide a histogram of the number of steps.

```{r}
dataSum <- aggregate(x = data[c("steps", "interval")],FUN=sum, by = list(Group.date = data$date))
meanSteps <- mean(dataSum$steps, na.rm=TRUE)
medianSteps <- median(dataSum$steps, na.rm=TRUE)
hist(dataSum$steps, xlab="Number of steps", ylab="Number of days", main="Number of steps taken per day")
```

The mean number of steps is `r meanSteps`, and the median number of steps taken per day is `r medianSteps`.

## Question 2: What is the average daily activity pattern?

Now we will make a time series plot of the average number of steps taken per 5-minute time interval, averaged over all days.

```{r}
dataSum2 <- aggregate(x = data[c("steps")],FUN=mean, by=list(Group.interval = data$interval), na.rm=TRUE)

plot(dataSum2$Group.interval, dataSum2$steps, type="l", ylim=c(0,200), xlab="Time interval", ylab="Average number of steps", main="Average number of steps by time interval")

maxInt = dataSum2[which.max(dataSum2$steps),1]
```

The 5-minute interval with the maximum number of steps is `r maxInt`.

## Question 3: Imputing missing values

There are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

```{r}
missingData <- nrow(data) - sum(complete.cases(data))
```

There are `r missingData` rows with missing values. Now, we will fill in the rows with missingData using the mean for that 5-minute interval. Then, we will recalculate the mean and median number of steps to see how our calculations have changed now that there are no missing values.

```{r}
data2 <- data
for (i in 1:nrow(data2)) {
  if (is.na(data2[i,1])) {
    count <- (data2[i,3]/5) + 1
    data2[i,1] <- dataSum2[count,2]
  }
}

newDataSum <- aggregate(x = data2[c("steps", "interval")],FUN=sum, by = list(Group.date = data$date))
newMeanSteps <- mean(newDataSum$steps, na.rm=TRUE)
newMedianSteps <- median(newDataSum$steps, na.rm=TRUE)
hist(newDataSum$steps, xlab="Number of steps", ylab="Number of days", main="Number of steps taken per day")
```

The new mean number of steps is `r newMeanSteps`, and the median number of steps taken per day is `r newMedianSteps`. By imputing the missing data, the estimates of the total daily number of steps has increased.

## Question 4: Are there differences in activity patterns between weekdays and weekends?

```{r}
data$date <- as.Date(data$date)
data[,4] <- weekdays(data$date)
if (data[,4] == "Monday") {
  data[,5] <- "weekday"
}
else if (data[,4] == "Tuesday"){
  data[,5] <- "weekday"
}
else if (data[,4] == "Wednesday"){
  data[,5] <- "weekday"
}
else if (data[,4] == "Thursday"){
  data[,5] <- "weekday"
}
else if (data[,4] == "Friday"){
  data[,5] <- "weekday"
}
else {
  data[,5] <- "weekend"
}

dataSum3 <- aggregate(x = data[c("steps", "V4", "V5")],FUN=mean, by=list(Group.interval = data$interval), na.rm=TRUE)

plot(dataSum3$Group.interval, dataSum2$steps, type="l", ylim=c(0,200), xlab="Time interval", ylab="Average number of steps", main="Average number of steps by time interval")
```

I didn't get to finish this part:

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.